Disease Knowledge Transfer across Neurodegenerative Diseases

### 2021-03-12 paper/pres/poster-dkt.tex
more on dkt poster
un-commited edits to poster

\usepackage[noend]{algorithm2e}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
\SetCommentSty{mycommfont}

@@

\newcommand{\lp}{\lambda_{d_i}^{k}}
\newcommand{\lpuu}{\lambda_{d_i}^{k,(u)}}
\newcommand{\lpum}{\lambda_{d_i}^{k,(u-1)}}

@@

\heading{Why}\;\;\;\;Posterior Cortical Atrophy (PCA): progression of \underline{multimodal} biomarkers not known $\rightarrow$ Identify outcome measures and suitable subjects for PCA clinical trials

@@

\begin{itemize}
 \item Diseases affect different brain regions un-equally, but underlyining mechanisms are the same (amyloid cascade)
 \item \textbf{Idea:} each brain region follows ``its own disease course'', common across diseases $\rightarrow$ dysfunction score
\end{itemize}

@@

\vspace{0.6em}
\begin{figure}[H]
 \centering
   \begin{subfigure}{0.49\columnwidth}
   \centering
   \fnt{1. Each disease characterised by region-specific dysfunction trajectories}\\
   $ \gamma_{ij}^l = f(\beta_{i} + m_{ij}; \lambda_{d_i}^l) $\\
%    \vspace{0.9em}
%    \includegraphics[width=\textwidth]{diseaseEvol}

  \end{subfigure}
  \begin{subfigure}{0.49\columnwidth}
   \centering
%    \vspace{-1.2em}
   \fnt{2. Dysfunction trajectory modelled using region-specific biomarkers}\\
   $ y_{ij}^k = g( \gamma_{ij}^{k} ; \theta_k) + N(0,\epsilon_k) $\\
%    \fnt{$y_{ijk}$ = measure in subject $i$, visit $j$, region $k$.}

@@

  \fnt{3. Extend to multiple subjects, biomarkers and diseases}\\
 $ p(\boldsymbol{y}|\theta, \lambda, \beta , \epsilon) = \prod_{(i,j,k) \in \Omega} p(y_{ij}^k|\theta_k, \lp, \beta_{i}) $
 \end{figure}

@@

\begin{figure}[H]
\begin{algorithm}[H]
\algoFnt{
%  Initialise $\boldsymbol{\theta}^{(0)}$, $\boldsymbol{\lambda}^{(0)}$, $\boldsymbol{\beta}^{(0)}$\\
  \While{$\boldsymbol{\theta}$, $\boldsymbol{\lambda}$, $\boldsymbol{\beta}$ not converged}{
   \tcp*[l]{\algoFnt{Estimate biomarker trajectories (disease agnostic)}}
%     \For{$k=1$ to $K$}{
      ${\theta_k\uu = \argmin_{\theta_k} \textcolor{red}{\sum_{(i,j) \in \Omega_k}} \left[y_{ijk} - g\left(f(\beta_i\um + m_{ij}; \lpum) ; \theta_k\right) \right]^2 }$\\
%       ${\epsilon_k\uu = \frac{1}{|\Omega_k|} \sum_{(i,j) \in \Omega_k}    \left[y_{ijk} - g\left(f(\beta_i\um + m_{ij}; \lpum) ; \theta_k\uu \right) \right]^2 }$\\
%     }
     \tcp*[l]{\algoFnt{Estimate dysfunction trajectories (disease specific)}} 
%     \For{$d=1 \in \mathbb{D}$}{
%       \For{$l=1 \in \Lambda$}{
        ${\lambda_{d}^{l, (u)} = \argmin_{\lambda_{d}^{l}} \textcolor{red}{\sum_{(i,j,k) \in \Omega_{d,l}}} \left[y_{ijk} - g\left(f(\beta_i\um + m_{ij}; \lambda_{d}^{l}) ; \theta_k\uu 
        \right) \right]^2 }$\\
%       }
%     }
    \tcp*[l]{\algoFnt{Estimate subject-specific time shifts}} 
%     \For{$i=1 \in [1, \dots, S]$}{
      ${\beta_i\uu = \argmin_{\beta_i} \textcolor{red}{\sum_{(j,k) \in \Omega_i}} \left[y_{ijk} - g\left(f(\beta_i + m_{ij}; \lpuu) ; \theta_k\uu
      \right) \right]^2 }$\\
%     }
}
}
% \normalfont
\end{algorithm}
% \caption[The algorithm for estimating the DKT parameters]{The algorithm used to estimate the DKT parameters, based on loopy belief-propagation.}
\label{fig:dktAlgo}
\end{figure}

@@

 \includegraphics[width=\columnwidth, trim=0 20 0 35, clip]{../figures/trajDisSpaceOverlap_PCA_tad-drcTinyPen5_JMD.png}
%  \caption{Estimated trajectories for the PCA cohort. The only data that were available were the MRI volumetric data. The dynamics of the other biomarkers has been inferred by the model using data from typical AD, and taking into account the different spatial distribution of pathology in PCA vs tAD. }
%  \label{fig:PCAtrajByModality}
\end{figure}

>>>>

\usepackage[noend]{algorithm2e}
\definecolor{comCol}{rgb}{0.0,0.0,0.0}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{comCol}{#1}}
\SetCommentSty{mycommfont}

@@

\definecolor{thetaCol}{rgb}{0.8,0.2,0.2}
\definecolor{lambdaCol}{rgb}{0.0,0.5,0.0}
\definecolor{betaCol}{rgb}{0,0.2,0.8}

\newcommand{\lp}{\textcolor{lambdaCol}{\lambda_{d_i}^{\psi(k)}}}
\newcommand{\lpuu}{\textcolor{lambdaCol}{\lambda_{d_i}^{\psi(k),(u)}}}
\newcommand{\lpum}{\textcolor{lambdaCol}{\lambda_{d_i}^{\psi(k),(u-1)}}}

@@

\heading{Why}\;\;\;\;Posterior Cortical Atrophy (PCA): progression of \underline{multimodal} biomarkers not known $\rightarrow$ Identify outcome measures and suitable subjects for PCA clinical trials.

@@

\begin{itemize}
 \item Diseases affect different brain regions un-equally, but underlyining mechanisms are the same (amyloid cascade).
 \item \textbf{Idea:} a brain region follows ``the same pathology course'' for both diseases, but to different extents and with different speed.
\end{itemize}

@@

\vspace{0.6em}
\begin{figure}[H]
 \centering
   \begin{subfigure}{0.49\columnwidth}
   \centering
   \fnt{1. Each disease characterised by region-specific \textcolor{lambdaCol}{dysfunction trajectories}}\\
   $ \gamma_{ij}^l = f(\textcolor{betaCol}{\beta_{i}} + m_{ij}; \textcolor{lambdaCol}{\lambda_{d_i}^l}) $\\
%    \vspace{0.9em}
%    \includegraphics[width=\textwidth]{diseaseEvol}

  \end{subfigure}
  \begin{subfigure}{0.49\columnwidth}
   \centering
%    \vspace{-1.2em}
   \fnt{2. Dysfunction trajectory modelled using region-specific \textcolor{thetaCol}{biomarker trajectories}}\\
   $ y_{ij}^k = g( \gamma_{ij}^{\psi(k)} ; \textcolor{thetaCol}{\theta_k}) + N(0,\epsilon_k) $\\
%    \fnt{$y_{ij}^k$ = measure in subject $i$, visit $j$, region $k$.}

@@

  \fnt{3. Extend to multiple subjects, biomarkers and diseases}\\
 $ p(\boldsymbol{y}|\textcolor{thetaCol}{\theta}, \textcolor{lambdaCol}{\lambda}, \textcolor{betaCol}{\beta}, \epsilon) = \prod_{(i,j,k) \in \Omega} p(y_{ij}^k|\textcolor{thetaCol}{\theta_k}, \lp, \textcolor{betaCol}{\beta_{i}}) $
 \end{figure}

@@

\begin{figure}[H]
\begin{algorithm}[H]
\algoFnt{
%  Initialise $\boldsymbol{\theta}^{(0)}$, $\boldsymbol{\lambda}^{(0)}$, $\boldsymbol{\beta}^{(0)}$\\
  \While{$\boldsymbol{\textcolor{thetaCol}{\theta}}$, $\boldsymbol{\textcolor{lambdaCol}{\lambda}}$, $\boldsymbol{\textcolor{betaCol}{\beta}}$ not converged}{
   \tcp*[l]{\algoFnt{Estimate \textcolor{thetaCol}{biomarker trajectories} (disease agnostic)}}
%     \For{$k=1$ to $K$}{
      ${\textcolor{thetaCol}{\theta_k\uu} = \argmin_{\textcolor{thetaCol}{\theta_k}} \textcolor{black}{\sum_{(i,j) \in \Omega_k}} \left[y_{ij}^k - g\left(f(\textcolor{betaCol}{\beta_i\um} + m_{ij}; \lpum) ; \textcolor{thetaCol}{\theta_k}\right) \right]^2 }$\\
%       ${\epsilon_k\uu = \frac{1}{|\Omega_k|} \sum_{(i,j) \in \Omega_k}    \left[y_{ij}^k - g\left(f(\beta_i\um + m_{ij}; \lpum) ; \theta_k\uu \right) \right]^2 }$\\
%     }
     \tcp*[l]{\algoFnt{Estimate \textcolor{lambdaCol}{dysfunction trajectories} (disease specific)}} 
%     \For{$d=1 \in \mathbb{D}$}{
%       \For{$l=1 \in \Lambda$}{
        ${\textcolor{lambdaCol}{\lambda_{d}^{l, (u)}} = \argmin_{\textcolor{lambdaCol}{\lambda_{d}^{l}}} \textcolor{black}{\sum_{(i,j,k) \in \Omega_{d,l}}} \left[y_{ij}^k - g\left(f(\textcolor{betaCol}{\beta_i\um} + m_{ij}; \textcolor{lambdaCol}{\lambda_{d}^{l}}) ; \textcolor{thetaCol}{\theta_k\uu} 
        \right) \right]^2 }$\\
%       }
%     }
    \tcp*[l]{\algoFnt{Estimate \textcolor{betaCol}{subject-specific time shifts}}} 
%     \For{$i=1 \in [1, \dots, S]$}{
      ${\textcolor{betaCol}{\beta_i\uu} = \argmin_{\textcolor{betaCol}{\beta_i}} \textcolor{black}{\sum_{(j,k) \in \Omega_i}} \left[y_{ij}^k - g\left(f(\textcolor{betaCol}{\beta_i} + m_{ij}; \lpuu) ; \textcolor{thetaCol}{\theta_k\uu}
      \right) \right]^2 }$\\
%     }
}
}
% \normalfont
\end{algorithm}
% \caption[The algorithm for estimating the DKT parameters]{The algorithm used to estimate the DKT parameters, based on loopy belief-propagation.}
\label{fig:dktAlgo}
\end{figure}

@@

%  \includegraphics[width=\columnwidth, trim=0 20 0 35, clip]{trajDisSpaceOverlap_PCA_tadDrcTiny_JMD}
  \includegraphics[width=\columnwidth, trim=0 20 0 35, clip]{../figures/trajDisSpaceOverlap_PCA_tad-drcTinyPen5_JMD.png}

%  \caption{Estimated trajectories for the PCA cohort. The only data that were available were the MRI volumetric data. The dynamics of the other biomarkers has been inferred by the model using data from typical AD, and taking into account the different spatial distribution of pathology in PCA vs tAD. }
%  \label{fig:PCAtrajByModality}
\end{figure}
